Supporting Information : Asymmetric tail-dependence in extreme climatic events affects population-extinction risk

### 2019-08-26 MainText.Rmd
deleting G-SG results, only include C-SC as moderate tail dep.

---
title: "Extreme climatic events with asymmetric tail-dependence affect population-extinction risk"
author: "Shyamolina Ghosh, Daniel Reuman"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=1in,right=1in,top=1in,bottom=1in"
output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty

@@

For single patch linear age-structured models, we found left-tail 
dependence (Clayton, Suvival Gumbel copulas) in noisy vital 
rates enhanced the extinction risk compared to the case when
vital rates were right tail dependent (Survival clayton, Gumbel copulas).
It was expected because a sudden "bad" year (e.g. with severe cold) can suppress 
the fecundity of adult and survival/ maturation of eggs simultaneously 
and thereby increasing the extinction risk of the population. For non-linear age-structured 
models, we got the similar trend for extinction risks with different tail-dependence (left and right).
With non-linear model the extinction risks appeared as higher than their corresponding 
linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG} represents this scenario with 4 different 
asymmetric copulas but of same Kendall's correlation ($\tau = 0.6$). There was very little
difference between extinction risks within a pair of different tail-dependent copulas
(e.g. C-SC in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG}(A,C) and
G-SG in Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG}(B,D)) and that difference was partly 
dependent on the strength of tail dependence. Therefore we repeated the same analysis with 
vital rates coming from extreme-tail dependent (left and right in two different simulations)
copulas and got the similar results but with much prominent difference between their extinction risks
(Fig. \ref{fig_extrisk_agestr}). This result answered (Q1) and showed extinction risk of 
a population could be considered as
direct manifestation of tail-dependence measure between vital rates of different age-class
within that population. <!--discussion later : add Tuljapurkar's approximation-->

@@

<!--Tuljapurkar approximation : linear version of age-structured model in stochastic environments-->
We found left tail dependence in vital rates (induced by ECEs) caused higher extinction risk than 
the right tail dependent ones, but the difference was not remarkable (Fig. \ref{SM-fig_extrisk_agestr_C_SC_G_SG}) 
unless the tail dependence was at its extreme (Fig. \ref{fig_extrisk_agestr}). This finding could be explained 
by the approximation @tuljapurkar1990lecture made for the stochastic log(growth rate) in an iid environment.
The approximation considered that the growth rate depends on the most dominant eigen value of the mean projection matrix 
and the temporal variance of the log(population growth rate) caused by environmental stochasticity, which was based
on the linear covariance measure between any two matrix elements, not on their tail-dependence properties. Therefore,
according to Tuljapurkar's approximation, population growth rate in an iid environment 
remained same for our model (Eq. \ref{eq_age_str}) with left-tail or right-tail dependent noises.

@@

 <!--FIG : age structure model results : linear, nonlinear-->
\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models
against time (shading represents 95\% confidence interval). Fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000. \label{fig_extrisk_agestr}}
\end{center}
\end{figure*}

<!--FIG :  summary : all nonlinear model results for local dispersion : extrisk vs. d-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp.pdf}
\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary} with local dispersal
among $n=5$ patches, extinction risk after $numsteps = 25$ years for $numsims = 10000$ simulations were calculated.
Extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$. Last two columns represent extinction risk calculated for intrinsic growth rate $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_localdisp}}
\end{center}
\end{figure*}

 <!--FIG : all nonlinear model results with d=0, extrisk vs. r-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend.pdf}\\
\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0.pdf}
\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0.pdf}
\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0.pdf}
\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0.pdf}\\
\caption{Transition of higher to lower extinction risk (calculated after $25$ years) for 
the metapopulations affected by extreme left tail dependent noise across
$5$ multiple patches without dispersal ($d=0$), as intrinsic 
growth rate ($r$) goes from under to over-compensatory regime. Opposite trend
was observed with extreme right tail dependent noise.
Error bars represent the standard errors associated with $10000$
simulations for extinction risk computed at each $r$. Dotted line represents the value of $r$
for each model where their deterministic analog switches from under- to over-compensatory dynamics.
\label{fig_extrisk_d0}}
\end{center}
\end{figure*}

>>>>

---
title: "Asymmetric tail-dependence in extreme climatic events affects population-extinction risk"
author: "Shyamolina Ghosh, Daniel Reuman"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=1in,right=1in,top=1in,bottom=1in"
output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty

@@

For single patch linear age-structured models, we found left-tail 
dependence (Clayton, Suvival Gumbel copulas) in noisy vital 
rates enhanced the extinction risk compared to the case when
vital rates were right tail dependent (Survival clayton, Gumbel copulas).
It was expected because a sudden "bad" year (e.g. with severe cold) can suppress 
the fecundity of adult and survival/ maturation of eggs simultaneously 
and thereby increasing the extinction risk of the population. For non-linear age-structured 
models, we got the similar trend for extinction risks with different tail-dependence (left and right).
With non-linear model the extinction risks appeared as higher than their corresponding 
linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
Fig. \ref{SM-fig_extrisk_agestr_mtd} represents this scenario with 4 different 
asymmetric copulas but of same Kendall's correlation ($\tau = 0.6$). There was very little
difference between extinction risks within a pair of different tail-dependent copulas
(e.g. C-SC in Fig. \ref{SM-fig_extrisk_agestr_mtd}(A,C) and
G-SG in Fig. \ref{SM-fig_extrisk_agestr_mtd}(B,D)) and that difference was partly 
dependent on the strength of tail dependence. Therefore we repeated the same analysis with 
vital rates coming from extreme-tail dependent (left and right in two different simulations)
copulas and got the similar results but with much prominent difference between their extinction risks
(Fig. \ref{fig_extrisk_agestr_etd}). This result answered (Q1) and showed extinction risk of 
a population could be considered as
direct manifestation of tail-dependence measure between vital rates of different age-class
within that population. <!--discussion later : add Tuljapurkar's approximation-->

@@

<!--Tuljapurkar approximation : linear version of age-structured model in stochastic environments-->
We found left tail dependence in vital rates (induced by ECEs) caused higher extinction risk than 
the right tail dependent ones, but the difference was not remarkable (Fig. \ref{SM-fig_extrisk_agestr_mtd}) 
unless the tail dependence was at its extreme (Fig. \ref{fig_extrisk_agestr_etd}). This finding could be explained 
by the approximation @tuljapurkar1990lecture made for the stochastic log(growth rate) in an iid environment.
The approximation considered that the growth rate depends on the most dominant eigen value of the mean projection matrix 
and the temporal variance of the log(population growth rate) caused by environmental stochasticity, which was based
on the linear covariance measure between any two matrix elements, not on their tail-dependence properties. Therefore,
according to Tuljapurkar's approximation, population growth rate in an iid environment 
remained same for our model (Eq. \ref{eq_age_str}) with left-tail or right-tail dependent noises.

@@

 <!--FIG : age structure model results : linear, nonlinear-->
\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models against time (shading represents 95\% confidence interval). Fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000. \label{fig_extrisk_agestr_etd}}
\end{center}
\end{figure*}

<!--FIG :  summary : all nonlinear model results for local dispersion : extrisk vs. d-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp_extremenoise.pdf}
\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary} with local dispersal
among $n=5$ patches, extinction risk after $numsteps = 25$ years for $numsims = 10000$ simulations were calculated.
Extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$. Last two columns represent extinction risk calculated for intrinsic growth rate $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_localdisp}}
\end{center}
\end{figure*}

 <!--FIG : all nonlinear model results with d=0, extrisk vs. r-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend.pdf}\\
\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_extremenoise.pdf}
\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0_extremenoise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0_extremenoise.pdf}
\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0_extremenoise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0_extremenoise.pdf}
\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_extremenoise.pdf}\\
\caption{Transition of higher to lower extinction risk (calculated after $25$ years) for 
the metapopulations affected by extreme left tail dependent noise across
$5$ multiple patches without dispersal ($d=0$), as intrinsic 
growth rate ($r$) goes from under to over-compensatory regime. Opposite trend
was observed with extreme right tail dependent noise.
Error bars represent the standard errors associated with $10000$
simulations for extinction risk computed at each $r$. Dotted line represents the value of $r$
for each model where their deterministic analog switches from under- to over-compensatory dynamics.
\label{fig_extrisk_d0}}
\end{center}
\end{figure*}

### 2019-08-26 SuppMat.Rmd
deleting G-SG results, only include C-SC as moderate tail dep.

---
title: "Supporting Information : Extreme climatic events with asymmetric tail-dependence affect population-extinction risk"
author: "Shyamolina Ghosh, Daniel Reuman"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=1in,right=1in,top=1in,bottom=1in"

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
library(abind)
source("mtime.R")
```

```{r sim_linear_C_SC_G_SG, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")

@@

SC_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(SC_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
#---------------------------------------------------------------------------------

# Now for gumbel

fam<-4
copG<-gumbelCopula(2)
parG<-iTau(copG,tau = ktau)
par<-parG

G_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(G_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))

#---------------------------------------------------------------------------------

# Now for survival gumbel : should be the same parameter as of gumbel if kendall's tau is the same

fam<-14
par<-parG

SG_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(SG_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
#---------------------------------------------------------------------------------
```

```{r fig_agestr_C_SC_G_SG_linear, echo=F,results="hide"}
xL<-C_age$extdf
xR<-SC_age$extdf
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., C","vital rates : RTdep., SC"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()

xL<-SG_age$extdf
xR<-G_age$extdf
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., SG","vital rates : RTdep., G"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()
```

```{r sim_nonlinear_C_SC_G_SG, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")

@@

SC_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(SC_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
#---------------------------------------------------------------------------------

# Now for gumbel

fam<-4
copG<-gumbelCopula(2)
parG<-iTau(copG,tau = ktau)
par<-parG

G_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(G_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))

#---------------------------------------------------------------------------------

# Now for survival gumbel : should be the same parameter as of gumbel if kendall's tau is the same

fam<-14
par<-parG

SG_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(SG_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
#---------------------------------------------------------------------------------
```

```{r fig_agestr_C_SC_G_SG_nonlinear, echo=F,results="hide"}
xL<-C_age$extdf
xR<-SC_age$extdf
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_",omg,".pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., C","vital rates : RTdep., SC"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()

xL<-SG_age$extdf[1:26,]
xR<-G_age$extdf[1:26,]
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_",omg,".pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., SG","vital rates : RTdep., G"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()
```

```{r sim_agestr_with_extremetail_linear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")
source("ExtremeTailDep.R")
numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-Inf

@@

```{r sim_agestr_with_extremetail_nonlinear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")
source("ExtremeTailDep.R")
numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-100
# For extreme right tail dep.
mr<-retd(n = (numsteps*numsims),d = 2,rl =1,mn=0,sdev=1)
cop<-pnorm(mr) # to make normal to uniformly distributed
 tempo<-paste("./Results/age_str_results/extreme_right_taildep_nonlinear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()
nlas_exR<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(nlas_exR,paste(resloc,"sim_age_str_extreme_right_taildep.RDS",sep=""))
 # For extreme left tail dep.
ml<-retd(n = (numsteps*numsims),d = 2,rl =-1,mn=0,sdev=1)
cop<-pnorm(ml) # to make normal to uniformly distributed
 tempo<-paste("./Results/age_str_results/extreme_left_taildep_nonlinear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()
nlas_exL<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(nlas_exL,paste(resloc,"sim_age_str_extreme_left_taildep.RDS",sep=""))
```

@@

<!--Supp. FIG :  age structure model results for C, SC, G, SG with linear model (\Omega = infinity)-->
\begin{figure}[!h]
\begin{center}
\textbf{ \hspace{0.5 cm} (A) \hspace{4.4 cm} (B)} \\
\vspace{-0.2 cm}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf}\\
\textbf{ \hspace{0.5 cm} (C) \hspace{4.4 cm} (D)} \\
\vspace{-0.2 cm}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_SG_G_nonlinear_omg_100.pdf}\\
\caption[Extinction risk from single patch age structured models with time when vital rates have either left (C, SG) or right tail (SC, G) dependence.]{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time; fecundity of adults and survival of eggs have either left or right tail dependence in their corresponding copula structure. $(A)$ Clayton and Survival Clayton (parameter = 3), $(B)$ Survival Gumbel and Gumbel (parameter = 2.5) for linear model ($\Omega = \infty$), numsims=10000. $(C - D)$ represents the analogous figures as of $(A - B)$ but for non-linear model ($\Omega = 100$). All copula has same Kendall correlation $\tau = 0.6$. \label{fig_extrisk_agestr_C_SC_G_SG}}
\end{center}
\end{figure}

>>>>

---
title: "Supporting Information : Asymmetric tail-dependence in extreme climatic events affects population-extinction risk"
author: "Shyamolina Ghosh, Daniel Reuman"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=1in,right=1in,top=1in,bottom=1in"

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
library(abind)
source("mtime.R")
# Note : for moderate tail dep. we used Kendall's tau = 0.6 to get C or SC
```

```{r sim_linear_mtd, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
# extinction risk with moderate tail dep. for single patch linear age structure model

set.seed(seed)
source("ext_risk_cop_age_str.R")

@@

SC_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(SC_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
```

```{r fig_agestr_mtd_linear, echo=F,results="hide"}
xL<-C_age$extdf
xR<-SC_age$extdf
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., C","vital rates : RTdep., SC"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()
```

```{r sim_nonlinear_mtd, echo=F,  cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"))}
# extinction risk with moderate tail dep. for single patch non-linear age structure model

set.seed(seed)
source("ext_risk_cop_age_str.R")

@@

SC_age<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(SC_age,paste(resloc,BiCopName(family = fam),"_sim_age_str.RDS",sep=""))
```

```{r fig_agestr_mtd_nonlinear, echo=F,results="hide"}
xL<-C_age$extdf
xR<-SC_age$extdf
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_",omg,".pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., C","vital rates : RTdep., SC"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()
```

```{r sim_agestr_with_extremetail_linear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
# extinction risk with extreme tail dep. for single patch linear age structure model

set.seed(seed)
source("ext_risk_cop_age_str.R")
source("ExtremeTailDep.R")
numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-Inf

@@

```{r sim_agestr_with_extremetail_nonlinear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
# extinction risk with extreme tail dep. for single patch non-linear age structure model

set.seed(seed)
source("ext_risk_cop_age_str.R")
source("ExtremeTailDep.R")
numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-100
# For extreme right tail dep.
mr<-retd(n = (numsteps*numsims),d = 2,rl =1,mn=0,sdev=1)
cop<-pnorm(mr) # to make normal to uniformly distributed
 tempo<-paste("./Results/age_str_results/extreme_right_taildep_nonlinear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()
nlas_exR<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(nlas_exR,paste(resloc,"sim_age_str_extreme_right_taildep.RDS",sep=""))
 # For extreme left tail dep.
ml<-retd(n = (numsteps*numsims),d = 2,rl =-1,mn=0,sdev=1)
cop<-pnorm(ml) # to make normal to uniformly distributed
 tempo<-paste("./Results/age_str_results/extreme_left_taildep_nonlinear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()
nlas_exL<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(nlas_exL,paste(resloc,"sim_age_str_extreme_left_taildep.RDS",sep=""))
```

@@

<!--Supp. FIG :  age structure model results for C, SC with linear model (\Omega = infinity)-->
\begin{figure}[!h]
\begin{center}
\textbf{ \hspace{0.5 cm} (A) \hspace{4.4 cm} (B)} \\
\vspace{-0.2 cm}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
\caption[Extinction risk from single patch age structured models with time when vital rates have either left (C) or right tail (SC) dependence.]{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time for 10000 simulations. Fecundity of adults and survival of eggs have moderate tail dependence in either left or right tail in their corresponding copula structure. Moderate left tail dependence (or right tail dependence) was generated by a bivariate Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6, $(A)$ for linear model ($\Omega = \infty$) and for $(B)$ non-linear model ($\Omega = 100$).\label{fig_extrisk_agestr_mtd}}
\end{center}
\end{figure}
